---
title: "NYPD Motor Vehicle Collisions Research Part#1"
author: "Glenda Ascencio"
date: "Started: Nov 12, 2015. Updated: October 11, 2016"
output: html_document
---

##### This research investigates the major causes that augmented the number of car accidents in NYC. More specifically, we’re comparing how the seasons, locations, years and months, and the Daylight Saving Time played a major role in augmenting the number of car collisions in NYC from 2012 to September 13, 2016. We obtained our data from the Motor Vehicle Collisions Department in New York City provided by the Police Department (NYPD) which has detailed information about more than 800,000 reported car collisions in NYC. 

##### Data set: On this data set we have information about the boroughs, locations, street names, total people injured, total people killed, total pedestrians injured, total pedestrians killed, total cyclist injured, total cyclist killed, total motorist injured, total motorist killed, reason vehicles crashed, and type of vehicle. All this data was imported, cleaned, and visualized into R

```{r echo=FALSE, message=FALSE, warning=FALSE}

##  Loading libraries
  #dm
library(dplyr)
  #dv
library(ggplot2)
library(plotly)
library(rbokeh)
library(scales)

##  Loading data
setwd("~/Github/Car_Collisions_Research")
load("car_collisions.RData")
```

##### First, I was trying to find the seasons (*Spring: March-May, *Summer: June-August, *Autumn: September-November *Winter: December-February) with the highest number of car collisions. According to our results, summer was the highest number of car collisions, followed by fall, spring, and lastly winter.

##### Season with highest car collision
```{r echo=FALSE, message=FALSE, warning=FALSE}
summary(car_collision$season)
seasons <- car_collision %>%
  group_by(season) %>%
  summarize(season_sum=n())

seasons <- arrange(seasons, desc(season_sum))
seasons$season <- factor(seasons$season,
                         levels=c("SUMMER", "FALL", "SPRING", "WINTER", "NA"),
                         ordered=TRUE)

plot_ly(seasons, x = ~season, y = ~season_sum, color = ~season, text = ~paste("people_killed: ", season_sum)) %>%
   layout(title = "2012-2016 Seasons With Highest Number of Car Collisions",
         scene = list(
           xaxis = list(title = "Seasons from 2012 to 2016"), 
           yaxis = list(title = "Total Number of Car Collisions")))
```


##### Later, we wanted to investigate in what borough we had the highest number of car collisions and our results show that Brooklyn is the most unsafe borough that has the major number of car collisions
```{r echo=FALSE, message=FALSE, warning=FALSE}
summary(car_collision$borough)

boroughs <- car_collision %>%
  group_by(borough) %>%
  summarize(borough_sum=n())

boroughs <- arrange(boroughs, desc(borough_sum))
boroughs <- boroughs[2:6,]

boroughs$borough <- factor(boroughs$borough, 
                   levels=c("BROOKLYN", "MANHATTAN", "QUEENS", "BRONX", "STATEN ISLAND"),
                   ordered=TRUE)

plot_ly(boroughs, x = ~borough, y = ~borough_sum, color = ~borough, text = ~paste("accidents: ", borough_sum)) %>%
   layout(title = "Boroughs with the highest number of car collisions",
         scene = list(
           xaxis = list(title = "Boroughs from 2012 to 2016"), 
           yaxis = list(title = "Total Number of Car Collisions")))

```

##### Investigating the day of the week is also an essential task we have to do know so we can educate our family about the day of the week that is the safest to travel

```{r}

car_collision <- mutate(car_collision, day_w = strftime(car_collision$date_time, format = "%A"))

#plot how many total accidents there are by day
day_accidents <- car_collision %>% group_by(day) %>% summarise(total_day = n())
day_accidents <- filter(day_accidents, day != "NA")
day_accidents <- arrange(day_accidents, desc(total_day))

day_accidents$day <- 
  factor(day_accidents$day, 
         levels = c("Friday", "Tuesday", "Thursday", "Wednesday", "Monday",
                    "Saturday", "Sunday"), ordered = T)

plot_ly(day_accidents, x = ~day, y = ~total_day, color = ~day, text = ~paste("t_accidents_p_day: ", total_day)) %>%
   layout(title = "Total accidents per day from 2012 to 2016",
         scene = list(
           xaxis = list(title = "Day of the week"), 
           yaxis = list(title = "Total Number of Car Collisions")))

```


##### We would also like to investigate some of the major causes of car accidents. According to our results driving inexperience/unattentitly, having fatigued or being drowsy, and not being able to yield right of way were the three major causes of car collisions.
```{r echo=FALSE, message=FALSE, warning=FALSE}
a_type <- car_collision %>% 
  group_by(year, season, reason_vehicle1_crashed) %>% 
  summarize(accidents = n())

a_type <- filter(a_type, reason_vehicle1_crashed !="Unspecified")
a_type <- arrange(a_type, desc(accidents))

a_type_highest <- filter(a_type, accidents >= 417)
a_type_lowest <- filter(a_type, accidents <= 416 & accidents >=20)

ggplot(a_type_highest, aes(x=year, y=accidents, fill=reason_vehicle1_crashed))+
  geom_bar(stat="identity")+
  ggtitle("Highest contribution factors that most car accidents occured >= 417") +
  xlab("2012 to 2016 Years")+
  ylab("Total Number of Car Collisions")


ggplot(a_type_lowest, aes(x=year, y=accidents, fill=reason_vehicle1_crashed))+
  geom_bar(stat="identity")+
  ggtitle("Lowest Reasons that most car accidents occured <= 416") +
  xlab("2012 to 2016 Years")+
  ylab("Total Number of Car Collisions")
```

##### In addition, I also wanted to investigate what were the most contribution factor that that caused more car accidents per season. 
```{r echo=FALSE, message=FALSE, warning=FALSE}
  # a) Calculations
reasons_type <- car_collision %>% 
  group_by(borough, year_month, season, reason_vehicle1_crashed) %>% 
  summarize(reasons = n())

  # b) Arranging the data so I can obtain the high reasons that caused car collisions
reasons_type <- filter(reasons_type, reason_vehicle1_crashed != "Unspecified")
reasons_type <- filter(reasons_type, reasons >= 100)

  # c) Plot the highest causes of car collisions 
ggplot(reasons_type, aes(x = reason_vehicle1_crashed, y = reasons,color = as.factor(borough)))+
  geom_point(stat="identity")+
  geom_point(aes(shape = as.factor(borough)))+
  scale_size_manual(values=c(3,3,3,3))+
  ggtitle("Reasons That Most Car Accidents Occured")+
  xlab("Reasons vehicle crashed")+
  ylab("Total Number of invididuals killed")+
  facet_wrap(~ season)+
  scale_y_continuous(limits=c(100, 600))+
  theme(legend.title=element_blank(),
        axis.text.x=element_text(angle=80, hjust=1),
        legend.background = element_rect(fill = "transparent"))
```


##### Furthermore, we wanted to investigate for each year, month, and borough, -What was the type of vehicle that had the most number of accidents?
```{r echo=FALSE, message=FALSE, warning=FALSE}
##  5) For each ym and borough, -What was the type of vehicle that had the most number of accidents?
  # a) Calculations
vehicle_type <- car_collision %>% 
  group_by(borough, year_month, reason_vehicle1_crashed, type_of_vehicle) %>% 
  summarize(vehicle_types = n())

  # b) Arranging the data so I can plot the highest vehicle class reasons and getting rid of unspefied
vehicle <- filter(vehicle_type, vehicle_types >= 30  & 
                    reason_vehicle1_crashed != "Unspecified")

  # Graphing the type of vehicle
ggplot(vehicle, aes(x = year_month, y = vehicle_types, colour = type_of_vehicle))+
  geom_point()+
  ggtitle("Car Accidents by Vehicle Facet_wrap by Borough")+
  xlab("Month of Year of the Accident")+
  ylab("Total Number of invididuals killed")+
  facet_wrap(~ borough)+
  scale_y_continuous(limits=c(10, 340))+
  theme(legend.title=element_blank(),
        axis.text.x=element_text(angle=80, hjust=1),
        legend.background = element_rect(fill = "transparent"))
```

##### We want citizens of NYC to know the satiest areas and dangerous areas in NYC so they can be aware were they should be more careful when traveling safety with their own family. That's why I also wanted to investigate in what year and in what location  pedestrian were injured the most?
```{r echo=FALSE, message=FALSE, warning=FALSE}
  # a) calculations
injured <- car_collision %>% 
  group_by(borough, season, year_month, total_people_injured) %>% 
  summarize(t_p_injured = n())

injured <- filter(injured, total_people_injured >=1)
injured <- arrange(injured, desc(t_p_injured))

View(injured)
  
# b) Plotting
ggplot(injured, aes(x = year_month, y = t_p_injured, color = as.factor(borough)))+
  geom_point(aes(shape = as.factor(borough), size = 2))+
  ggtitle("Perilous location where pedestrian have been injured the most")+
  xlab("Year and Month")+
  ylab("Total Number of People Injured")+
  theme(legend.title=element_blank(),
        axis.text.x=element_text(angle=80, hjust=1),
        legend.background = element_rect(fill = "transparent"))

# For each borough how many people were injured?
injured1 <- car_collision %>%
  group_by(borough, total_people_injured)%>%
  summarize(t_inj=n())
injured1 <- filter(injured1, total_people_injured != "0" & borough != "")
injured1 <- arrange(injured1, desc(t_inj))

figure(width = NULL, height = NULL, 
       title = "Total people injured in each borough",
       xlab = "Total people injured)",
       ylab = "The sum of total people injured in each borough") %>%
  ly_points(total_people_injured, t_inj, data = injured1, 
    color = borough, glyph = borough,
    hover = list(total_people_injured, t_inj))

```


##### In NYC we also have a plethora of motorist who like to travel anywhere in NYC. However, it would be good to informed them about the most perilous boroughs, year and months where more motorist accidents have happened the most so they could be travel more cautiously
```{r echo=FALSE, message=FALSE, warning=FALSE}
  # a) Calculations
motorist_accidents <- car_collision %>% 
  group_by(borough, season, year_month, total_motorist_killed) %>% 
  summarize(n_motorist_killed = n())

motorist_killed <- filter(motorist_accidents, total_motorist_killed == "1")
motorist_killed <- arrange(motorist_killed, desc(n_motorist_killed))

head(motorist_killed, n=10)

  # b) Plotting
ggplot(motorist_killed, aes(x = year_month, y = n_motorist_killed, color = as.factor(year_month)))+
  geom_point(aes(shape = as.factor(season), size = .2))+
  ggtitle("Boroughs where most motorist were killed the most")+
  xlab("Year and month")+
  ylab("Total number of motorist killed")+
  facet_wrap(~ borough) +
  theme(legend.title=element_blank(),
        axis.text.x=element_text(angle=80, hjust=1),
        legend.background = element_rect(fill = "transparent"))


ggplot(motorist_killed, aes(x = year_month, y = n_motorist_killed, color = as.factor(season)))+
   geom_point(aes(shape = as.factor(season), size = .2))+
  ggtitle("Boroughs Where Most Motorist Were Killed The Most")+
  xlab("Year and Month color coded by season")+
  ylab("Total Number of Motorist Killed")+
  facet_wrap(~ borough) +
  theme(legend.title=element_blank(),
        axis.text.x=element_text(angle=80, hjust=1),
        legend.background = element_rect(fill = "transparent"))

m_k <- motorist_killed %>%
  group_by(season) %>%
  summarise(total=n())

plot_ly(m_k, x = ~season, y = ~total, color = ~season, text = ~paste("people_killed: ", total)) %>%
   layout(title = "Seasons Where Most Motorist Were Killed The Most",
         scene = list(
           xaxis = list(title = "Seasons"), 
           yaxis = list(title = "Number of Motorist Killed")))

```



##### We would also like to know in what month we had the borough with the highest number of car collision. Our results show that month of July (20094) and August (19730) had the greatest number  of car collisions on Brooklyn. Similarly, on the month of July we have the boroughs of Manhattan with 19730 and the borough of Queens with 17220 number of car accidents. 
```{r echo=FALSE, message=FALSE, warning=FALSE}
b_ym <- car_collision %>%
  group_by(month, borough) %>%
  summarize(b_ym_sum=n())

b_ym <- arrange(b_ym, desc(b_ym_sum))

b_ym <-  filter(b_ym, 
                  !is.na(month) & 
                  !is.na(borough) &
                  !month == "" &
                  !borough == "")

ggplot(b_ym, aes(x=month, y=b_ym_sum, fill=borough)) +
  geom_bar(stat="identity") +
  geom_smooth(se = FALSE, method = "loess") +
  ggtitle("12 Months with highest number of car collisions from 2012-2016") +
  xlab("Months")+
  ylab("Total Number of Car Collisions")

```

##### We would like to further investigate why in the month of July we have the highest number of car collisions on the boroughs of Brooklyn, Manhattan, and Queens. That's why we would take a look to the month of July from the year 2012 to 2016 to see which month had the highest number of car collisions. According to our results, July 2015 had the highest number of car collisions and this might be caused by the more than 30 holidays that we celebrate in the United States such as the USA Independence Day, Parent's Day, and the International Joke Day. 

```{r echo=FALSE, warning=FALSE, message=FALSE}
#1
july <- car_collision %>%
  group_by(year_month, borough) %>%
  summarize(july_total=n())

july <- filter(july, 
              (year_month == "2012-07" |
                year_month == "2013-07" |
                year_month == "2014-07" |
                year_month == "2015-07" |
                year_month == "2016-07") & 
              (!borough == ""))
july <- arrange(july, desc(july_total))
july$year_month <- factor(july$year_month,
                         levels=c("2015-07", "2013-07", "2014-07", "2012-07", "2016-07"),
                         ordered=TRUE)

#2
julys <- car_collision %>%
  group_by(date, year_month, borough) %>%
  summarize(julys_total=n())

julys1 <- filter(julys, 
              (year_month == "2012-07" |
                year_month == "2013-07" |
                year_month == "2014-07" |
                year_month == "2015-07" |
                year_month == "2016-07") & 
              (!borough == ""))
julys1 <- arrange(julys1, desc(julys_total))
head(julys1)

print("This first graph displays all the months of July from the year 2012 to 2016. This graph, displays the total number of car collision as a function of month and year. Our results show that on July 2015 we had the highest number of car collisions")
ggplot(july, aes(x=year_month, y=july_total, fill=year_month)) +
  geom_bar(stat="identity") +
  geom_smooth(se = FALSE, method = "loess") +
  ggtitle("Comparing each July months from 2012 to 2016") +
  xlab("Months")+
  ylab("Total Number of Car Collisions")

print("This second graph displays the total number of car collision as a function of months color coded by borough. In this graph we can see how Brooklyn has the highest number in car collisions on July 2014")
 ggplot(na.omit(julys1), aes(x = year_month, y = julys_total, colour=borough)) + 
  geom_boxplot(aes(fill = borough), alpha = 0.1) + 
   guides(fill=FALSE) + 
  labs(x = "Months", 
       y = "Total Number of Car Collisions", 
       title = "July months with highest number of car collisions by borough") 

print("This third graph displays the total number of car collision as a function of dates color coded by borough. This graph shows that on July 3 2014 we have highest number of car collisions on Brooklyn")
p <- ggplot(julys1, aes(x=date, y=julys_total, group=borough, colour=borough))
p + geom_line(arrow = arrow())+
   geom_point(aes(shape = as.factor(borough), size = .2))+
  ggtitle("July months with highest number of car collisions by borough") +
  xlab("Dates")+
  ylab("Total Number of Car Collisions")+
  theme(legend.title=element_blank(),
        axis.text.x=element_text(angle=90, hjust=1),
        legend.background = element_rect(fill = "transparent")) 
```

##### In conclusion, thousands of individuals ages from 1 to 90 are being killed in each borough due to driving inattention (dictraction due the use of cellphones, putting make up), failure to yield right of way, fatigue, and driving in a drunk condition. Many of these accidents occur during the workday or during the commute to and from work/home. Our results show that the summer is the most perilous season in NYC. Citizens ought to drive carefully in Brooklyn and Manhattan because those were the boroughs with the most car accidents from the years 2012 to 2016. Our results also demonstrate that the majority of car collision occur on passenger vehicles which means that a plethora of inviduals are using their personal cars for transportation. Driving innattentively, being tired or drowsy, failure to yield-right-away, and backing unsafely are the biggest factors for contributing and augmenting the number of car collisions. That's why in order to decrease the number of car collisions we need to drive carefully, pay attention to our surrunding areas, and remain patitient at all times.
